---
title: "R Notebook"
output: html_notebook
---

```{r}
# install.packages("palmerpenguins")
library(palmerpenguins)
library(tidyverse)

penguins <- palmerpenguins::penguins 
```

```{r}
# three species of penguin
penguins %>% 
  distinct(species)
```
EXPLORATORY ANALYSIS

```{r}
ggplot(penguins)+
  aes(x = species, y = bill_length_mm)+
  geom_boxplot()+
  labs(x = "Species",
       y = "Bill Lenghth (mm)")
```
Each box represents 50% of data per specific species. 
The horizontal line represents the middle value of the stored data set.
The whiskers on either side of the box mark the positions beyond which data values are normally deemed to be outliers.
In our case there is only one outlier in Gentoo species.
In each case, we can see fair distribution around the median.

```{r}
penguins %>%
  drop_na(bill_length_mm) %>% 
  filter(species == "Chinstrap") %>% 
  ggplot(aes(x = bill_length_mm)) +
  geom_histogram(col = "white")
```
It’s difficult to say clearly, but Chinstrap's bill length distribution shows some evidence of being bimodal.

```{r}
penguins %>%
  drop_na(bill_length_mm) %>% 
  filter(species == "Gentoo") %>% 
  ggplot(aes(x = bill_length_mm)) +
  geom_histogram(col = "white")
```
Again, it isn't easy to say clearly, but Gentoo's bill length distribution looks to be slightly right skewed.


```{r}
penguins %>%
  drop_na(bill_length_mm) %>% 
  filter(species == "Adelie") %>% 
  ggplot(aes(x = bill_length_mm)) +
  geom_histogram(col = "white")
```
In the case of the Adelie species bill length, we can see normal distribution but there isn't a clear peak in the middle. 

```{r}
penguins %>%
  drop_na(bill_length_mm) %>% 
  group_by(species) %>%
  summarise(IQR = IQR(bill_length_mm))

penguins %>%
  drop_na(bill_length_mm) %>% 
  group_by(species) %>%
  summarise(
    Q1 = quantile(bill_length_mm, 0.25),
    Q2 = quantile(bill_length_mm, 0.5),
    Q3 = quantile(bill_length_mm, 0.75),
    IQR = Q3 - Q1
  )
```

```{r}
ggplot(penguins) +
  geom_pointrange(aes(x = species, y = bill_length_mm),
                  stat = "summary",
                  fun.min = function(z) {quantile(z,0.25)},
                  fun.max = function(z) {quantile(z,0.75)},
                  fun = median)
```

```{r}
ggplot(data = penguins, aes(y = bill_length_mm, x = species)) + # Move y and x here so than they can be used in stat_*
    geom_dotplot(aes(fill = species),   # Use fill = Species here not in ggplot()
                 binaxis = "y",         # which axis to bin along
                 binwidth = 0.6,        # Minimal difference considered diffeerent
                 stackdir = "center"    # Centered
                 ) +
    stat_summary(fun = median, fun.min = median, fun.max = median,
                 geom = "crossbar", width = 0.5)+
  labs(x = "",
       y = "Bill Lenghth (mm)",
       fill = "Species")
```


SUMMARY STATISTICS

--- Mean & Median --- 
```{r}
penguins %>% 
  drop_na(bill_length_mm) %>% 
  group_by(species) %>% 
  summarise(mean = mean(bill_length_mm), 
            median = median(bill_length_mm),
            difference = mean(bill_length_mm) - median(bill_length_mm))
```
Average length of the bill between penguin species (Adelie, Chinstrap, Gentoo) is 38.79 mm, 48.83 mm, 47.50 mm respectively.

Median length of the bill between penguin species (Adelie, Chinstrap, Gentoo) is 38.80 mm, 49.55 mm, 47.30 mm respectively.

Median and mean values of the bill length hardly differ for Adelie penguin species. 

The most significant difference, from the three species, in the average and median length of the bill is seen in Chinstraps (the numbers vary by 0.72 mm).


--- Spread (Standard Deviation) ---
```{r}
penguins %>%
  drop_na(bill_length_mm) %>% 
  group_by(species) %>%
  summarise(standard_deviation = sd(bill_length_mm))
```
The standard deviation shows that the observations are most spread out for Chinstrap penguin.


--- Skewness ---
```{r}
library(e1071)

penguins %>%
  drop_na(bill_length_mm) %>% 
  group_by(species) %>%
  summarise(skewness = skewness(bill_length_mm, type = 1))
```
The distribution is fairly symmetrical for Adelie and Chinstrap.
And the distribution is moderately right-skewed for Gentoo.
