---
title: "Correlation matrix"
output: html_notebook
---

```{r}
library(palmerpenguins)
library(tidyverse)
library(GGally)

penguins <- palmerpenguins::penguins 
```

Let's create a correlation matrix using GGally::ggcorr() that shows the correlation between bill length, bill depth, flipper length, and body mass across species.

```{r}
penguins <- penguins %>% 
  select(c(bill_length_mm, bill_depth_mm, flipper_length_mm, body_mass_g)) %>% 
  drop_na()
```

```{r}
ggcorr(penguins, nbreaks = 7,palette = "RdGy", label = TRUE, label_size = 3, label_color = "white")
```
**flipper length vs body mass**
```{r}
ggplot(penguins)+
  aes(x = flipper_length_mm, y = body_mass_g)+
  geom_point()
```
Positively correlated. Very strong correlation.

**flipper length vs bill depth**
```{r}
ggplot(penguins)+
  aes(x = flipper_length_mm, y = bill_depth_mm)+
  geom_point()
```
Not correlated.

**body mass vs bill depth**
```{r}
ggplot(penguins)+
  aes(x = body_mass_g, y = bill_depth_mm)+
  geom_point()
```
Not correlated.

**body mass vs bill length**
```{r}
ggplot(penguins)+
  aes(x = body_mass_g, y = bill_length_mm)+
  geom_point()
```
Positively correlated. Strong correlation.

**flipper length vs bill length**
```{r}
ggplot(penguins)+
  aes(x = flipper_length_mm, y = bill_length_mm)+
  geom_point()
```
Positively correlated. Strong correlation.

**bill depth vs bill length**
```{r}
ggplot(penguins)+
  aes(x = bill_depth_mm, y = bill_length_mm)+
  geom_point()
```
Not correlated.

MAIN FINDINGS

Across penguin species, there is a strong positive correlation between body mass and flipper length. 
There is a strong positive correlation between body mass and bill length. 
And also a strong positive correlation between flipper length and bill length.
Surprisingly, there is no correlation between body mass and bill depth or between bill length and bill depth.
Based on this, we can assume that heavier/larger penguins have longer bills and flippers.