---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```

1. Read the cleaned data into R.
```{r}
meteorites <- read_csv("data/meteorites_landings_clean.csv")
```

2. Find the names and years found for the 10 largest meteorites in the data.
```{r}
meteorites %>% 
  arrange(desc(mass_g)) %>% 
  select(name, year) %>% 
  drop_na() %>% 
  head(10)
```
3. Find the average mass of meteorites that were recorded falling, vs. those which were just found.
```{r}
meteorites %>% 
  group_by(fall) %>% 
  summarise(avg = mean(mass_g))
```

4. Looking for the number of meteorites in each year, for every year since 2000.
```{r}
meteorites %>% 
  filter(year >= 2000) %>% 
  group_by(year) %>% 
  summarise(count = n())
```
5.Write a function called get_meteorite_count() that takes in two arguments: a dataframe and a year, and returns the phrase “In (input_year), there were (number of meteorites that year) meteorites.” IF there are 1 or more meteorites that year. If there are no meteorites in the year, return the phrase “As far as we know, there were no meteorites that year”.

3.5 Writing function/program to perform statistical analysis

```{r}
get_meteorite_count <- function(dataframe, year){ # functions takes two arguments
  count = 0 # create a variable 
  if(year %in% dataframe$year){ # checks if input year is in a dataset
    for(i in dataframe$year){ # loops through the rows in a year column
      if(!is.na(i)){ # excludes missing values
        if (i == year){ # if loop value equal to the selected
          count = count + 1 # add 1 to the variable count
        }
      }
    }
    print(paste0("In ", year, " , there were ", count, " meteorites.")) # prints selected year and the count
    return(TRUE) # returns True
  }
  else{
    print("As far as we know, there were no meteorites that year") # prints a message
    return(FALSE) # returns False
  }
}

get_meteorite_count(meteorites, 1804) # run the function
```
6. Plans on how the function could be improved, how could I test that the function works.


4.1 Test plans

TEST PLANS

--- Function Improvements ---

1. The function can have additional functionality by searching maximun and minimum year in the file and if a year was entered out of a scope, we can we informed about the scope in else statement.

2. The function could be improved by giving an error if a logical or character value was entered. 

--- Test Plans ---

We can use 'testthat' package to perform a unit testing. From where we can use two useful functions 'expect_equal' to check that the objects give us an answer/value we would expect, and 'expect_error' to check that code throws an error.



7.Write a unit test for your function to check whether non-numeric inputs (character and logical) for the year argument return an error.

4.6 Performing test and revising code

```{r}
library(testthat)

test_that("Non-numeric input returns an error", { # failure as it doesn't throw an error - and we expect one
  expect_error(get_meteorite_count(meteorites, "year"))
  expect_error(get_meteorite_count(meteorites, TRUE))
})
test_that("Meteorite count function includes", {
  expect_equal(get_meteorite_count(meteorites, 2020), FALSE)
  expect_equal(get_meteorite_count(meteorites, 2005), TRUE)
})
```
Improving the function based on the result from the unit test. Testing the function to make sure it now gives an error when we input a logical or character value as the year.
```{r}
refactor_get_meteorite_count <- function(dataframe, year){
  if (!is.numeric(year)){
    stop("Non-numeric input.")
  }
  count = 0
  if(year %in% dataframe$year){
    for(i in dataframe$year){
      if(!is.na(i)){
        if(i == year){
          count = count + 1
        }
      }
    }
    print(paste0("In ", year, " , there were ", count, " meteorites.")) 
    return(TRUE)
  }
  else{
    print("As far as we know, there were no meteorites that year")
    return(FALSE)
  }
}
test_that("Non-numeric input returns an error", {
  expect_error(refactor_get_meteorite_count(meteorites, "year"))
  expect_error(refactor_get_meteorite_count(meteorites, TRUE))
})
```













