---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
meteorites <- read_csv("data/meteorites_landings_clean.csv")
```


```{r}
meteorites %>% 
  ggplot()+
  aes(x = fall)+
  geom_bar()

meteorites %>% 
  drop_na() %>% 
  summarise(min = min(year),max = max(year))
```
3.6 Writing function/program to create visualisations
```{r}
meteorites %>% 
  ggplot()+
  aes(x = year, fill = fall)+
  geom_histogram(binwidth = 100, position=position_dodge())+
  labs(x = "Year",
       y = "Count",
       fill = "Status",
       title = "Meteorites fallen to Earth vs found up to 2013")
```

The histogram displays meteorites that have fallen to Earth versus 
those that have been found between the years 1600 up to 2013.

Data are grouped into buckets of equal width, one bucket represents 100 years.
Each bucket is plotted as a bar whose height corresponds to how many meteorites 
were detected in that 100 years depending on their status (fell vs found).

The graph is skewed left. On the left side, before the year 1800, 
there weren't any meteorites found despite that some were recorded as fallen.
Most of the meteorites were recorded found after the year 2000.

In regards to the numbers of meteorites that were fallen, 
the largest number was recorded to be between 1900 to 2000. 


```{r}
#ggsave("3_6_plot.png")
```





